---
title: "example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The example relies on `microclimr` for microclimate data analyses, 
`dplyr`, `tidyr` and `broom` for table manipulations, 
`lubridate` for dates wrangling, 
and `ggplot2` for graphic creations 
(a few will be replaced by the package functions itself).

```{r setup}
#| message: false
library(microclimr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(broom)
```

## Data

Data are composed of air temperature at 1-m above the ground at every hour for 
the HOBO sensor "59-32-a" in 2023 within plot "59-32" from the Mormal forest 
joined by date and time with ERA5-Land temperature at 2m been extracted from 
the Google Earth Engine (GEE) for longitude 3.70 and latitude 50.2 
corresponding to the Mormal forest.

```{r data}
data <- hobo |>
  left_join(rename(era, datetime = time), by = join_by(datetime)) |>
  mutate(season = ifelse(month_num %in% 5:11, "leaf-on", "leaf-off"))
data |>
  ggplot(aes(x = datetime)) +
  geom_line(aes(y = tas, col = "ERA5-Land"), alpha = .5) +
  geom_line(aes(y = t_hobo, col = "HOBO"), alpha = .5) +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Temperature [°C]") +
  xlab("") +
  scale_color_manual("", values = c("#e4c284", "#4a8b76")) +
  geom_vline(
    xintercept = as_datetime("2023-05-01 00:00:00"),
    linetype = "dashed", col = "darkgrey"
  )
```

## Offset

A simple initial assessment of the microclimate effect can be made by computing
the offset between the macro- and microclimatic temperatures by subtraction, as
described in De Frenne et al. (2019). At Mormal, we observe a negligible mean
offset during the leaf-off season, but a drop of approximately 1.1°C during the
leaf-on season.

```{r offset_fig}
data |>
  ggplot(aes(season, t_hobo - tas, col = season)) +
  geom_boxplot() +
  theme_bw() +
  scale_color_manual("Period", values = c("#816768", "#68ae72")) +
  ylab("Temperature offset [°C]") +
  xlab("") +
  theme(
    axis.text.x = element_blank(), axis.ticks.x = element_blank(),
    legend.position = "bottom"
  )
```

```{r offset_tab}
data |>
  group_by(season) |>
  summarise(offset = mean(t_hobo - tas)) |>
  knitr::kable()
```

## Slope and equilibrium

Another method, suggested by Gril et al. (2023), involves using a regression
between macroclimatic and microclimatic temperature to compute the slope of the
microclimate effect and the equilibrium temperature. This method is known as the
slope and equilibrium approach.  At Mormal, we observe a negligible slope effect
during the leaf-off season, but during the leaf-on season, we observe a slope
that reduces high temperature and increases low temperature in the understorey,
with a value of 0.7 and an equilibrium point at 12.6°C.

```{r se_fig}
data |>
  ggplot(aes(tas, t_hobo, col = season)) +
  geom_point(alpha = .01) +
  geom_abline(linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  scale_color_manual("Period", values = c("#816768", "#68ae72")) +
  xlab("ERA5-Land temperature [°C]") +
  ylab("HOBO temperature [°C]") +
  coord_equal() +
  theme(legend.position = "bottom")
```

```{r se_tab}
data |>
  group_by(season) |>
  do(fit = lm(t_hobo ~ tas, data = .)) |>
  mutate(fit = list(tidy(fit))) |>
  unnest() |>
  mutate(term = recode(term,
    "(Intercept)" = "intercept",
    "tas" = "slope"
  )) |>
  select(season, term, estimate) |>
  pivot_wider(names_from = "term", values_from = "estimate") |>
  mutate(equilibrium = intercept / (1 - slope)) |>
  select(season, slope, equilibrium) |>
  knitr::kable()
```

## Fourier Transform

The `microclimr` package proposes a new method based on the Fourier transform.
Short description [link to maths vignette].

First, we can quickly decompose the Fourier coefficients using a window of five
days sliding on three days by season (leaf-off or leaf-on) and source of data
(ERA macroclimate or HOBO microclimate), by using the `fft_roll` function. This
requires further explanation.

```{r fft_roll}
fft_all <- data |>
  select(season, datetime, tas, t_hobo) |>
  rename(era = tas, hobo = t_hobo) |>
  pivot_longer(era:hobo, names_to = "source", values_to = "temperature") |>
  group_by(source, season) |>
  do(fft = fft_roll(., 24 * 5, "datetime", "temperature")) |>
  unnest(fft)
```

Then, using the mean represented here by the 0 period, we can identify the
linear relationship between mean temperatures.

```{r fft_0}
fft_all |>
  filter(period == 0) |>
  select(season, source, datetime, power) |>
  pivot_wider(names_from = source, values_from = power) |>
  ggplot(aes(era, hobo, col = season)) +
  geom_point() +
  geom_smooth(method = "lm", formula = " y ~ x") +
  theme_bw() +
  scale_color_manual("Period", values = c("#816768", "#68ae72")) +
  xlab("ERA5-Land temperature [°C]") +
  ylab("HOBO temperature [°C]") +
  coord_equal() +
  theme(legend.position = "bottom") +
  ggtitle("Mean temperature")
```

We can then calculate the energy of all the coefficients to determine the amount
of energy dissipated during the leaf-on and leaf-off periods. This highlights
the microclimate effect during the leaf-on season, with 87% of the energy
remaining in the microclimate.

```{r fft_enery_fig}
fft_all |>
  group_by(source, season, datetime) |>
  summarise(energy = fft_energy(fft_coefficient(amplitude, phase))) |>
  pivot_wider(names_from = source, values_from = energy) |>
  ggplot(aes(season, hobo / era, fill = season)) +
  geom_boxplot() +
  theme_bw() +
  scale_fill_manual(guide = "none", values = c("#816768", "#68ae72")) +
  ylab("") +
  xlab("") +
  coord_equal() +
  theme(legend.position = "bottom") +
  ggtitle("Energy HOBO / ERA5-Land")
```

```{r fft_enery_tab}
fft_all |>
  group_by(source, season, datetime) |>
  summarise(energy = fft_energy(fft_coefficient(amplitude, phase))) |>
  pivot_wider(names_from = source, values_from = energy) |>
  group_by(season) |>
  summarise(dissipation = mean(hobo / era)) |>
  knitr::kable()
```

## Additional Fourier analyses

We can also represent the power spectrum of both the macro- and microlimates
after applying the Fourier transform (a graphical function to be added).

```{r power_spectrum}
fft_all |>
  filter(season == "leaf-on") |>
  group_by(source) |>
  filter(period != 0) |>
  ggplot(aes(frequency, power, fill = source)) +
  geom_col(position = "dodge") +
  theme_bw() +
  ggtitle("Representation of power spectrum (window 5days)") +
  xlab("Period [h]") +
  scale_x_continuous(
    breaks = c(0, 1 / 24, 1 / 12, 1 / 8, 1 / 6, 1 / 3),
    labels = c("0", "24", "12", "8", "6", "3")
  ) +
  scale_fill_manual("", values = c("#e4c284", "#4a8b76")) +
  theme(legend.position.inside = c(0.8, 0.8))
```

We can also represent the temperature amplitude relation for a specific
frequency, such as the 24-hour period shown here (graphical function to be
added?).

```{r ft_24}
fft_all |>
  filter(period == 24) |>
  select(season, source, datetime, power) |>
  pivot_wider(names_from = source, values_from = power) |>
  ggplot(aes(era, hobo, col = season)) +
  geom_point() +
  geom_smooth(method = "lm", formula = " y ~ x") +
  theme_bw() +
  scale_color_manual("Period", values = c("#816768", "#68ae72")) +
  xlab("ERA5-Land temperature [°C]") +
  ylab("HOBO temperature [°C]") +
  coord_equal() +
  theme(legend.position = "bottom") +
  ggtitle("Frequency 24h")
```

## Reconstruction

The power spectrum can also be used to reconstruct time series with a finer time
step. For example, this can be achieved using the `fft_reconstruct` function (to
be used for macroclimate debiasing).

```{r fft_reconstruct}
sub <- filter(
  fft_all, source == "hobo",
  datetime == as_datetime("2023-01-04 12:30:00")
)
tibble(time = seq(0, 24 * 5, by = 0.1)) |>
  mutate(temperature = fft_reconstruct(
    coefficients = fft_coefficient(sub$amplitude, sub$phase),
    frequencies = sub$frequency[-1],
    time = time
  )) |>
  ggplot(aes(
    as_datetime(as_date("2023-01-04") - 2 + (time + 1) / 24),
    temperature
  )) +
  geom_line() +
  theme_bw() +
  ylab("Temperature [°C]") +
  xlab("Day") +
  geom_point(
    data = filter(
      data,
      datetime >= as_date("2023-01-02"),
      datetime <= as_date("2023-01-07")
    ),
    aes(x = datetime, y = t_hobo), col = "red"
  )
```


## References

De Frenne, P., Zellweger, F., Rodríguez-Sánchez, F. et al. Global buffering of
temperatures under forest canopies. Nat Ecol Evol 3, 744–749 (2019).
https://doi.org/10.1038/s41559-019-0842-1

Gril, E., Spicher, F., Greiser, C., Ashcroft, M. B., Pincebourde, S., Durrieu,
S., ... & Lenoir, J. (2023). Slope and equilibrium: A parsimonious and flexible
approach to model microclimate. Methods in Ecology and Evolution, 14(3),
885-897.  https://doi.org/10.1111/2041-210X.14048
